---
title: "HW2OliviaD'Souza"
author: "Olivia D'Souza"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 
Work on Lab 1 if you haven’t done so already. Be sure to read the instructions and the
sample submissions carefully before you start. You are expected to reproduce most of
the outputs in Lab 1 by the end of class on Tuesday, Sept 12.

## Question 2
(This will be in Quiz.) Please review the concept of computing p-values 
and think about how to find the p-value of a one-sided (aka. one-tailed) test from t
he p-value of a two-sided (aka. two-tailed) test. Here are some references:

The following questions use the Copier Maintenance data from Problem 1.20 (p.24)
## Question 3: Problem 2.5 (a, b, c, d) (2 pts)

```{r}
#read Copier Maintenance data in
cm <- read.table("ch2dataset.txt", header=F)
colnames(cm) <- c("time", "copiers")
head(cm, 3)
tail(cm, 2)
```


# a. Estimate the change in the mean service time when the number of copiers serviced increases by one. Use a 90 percent confidence interval. Interpret your confidence interval.Part (a) is asking about the slope. (Why?) 


```{r}

cm.SLR<-lm(time~copiers, data=cm) #estimated linear regression
cm.SLR
summary(cm.SLR) # In order to find the estimated parameter & standard error
critVal <-qt(.95,43) #finds critical value based on 90% confidence int
critVal
confint(cm.SLR, level=0.90) # confirms my math using R software


```

Part A is asking about the slope because it asks for the "change in mean service time (y variable)
when the number of copiers serviced (x variable) increases by one. This is the definition
of the slope since you interpret the value as the change in mean y for one unit increase of x. 

Confidence Interval = est. of b1 ± (critical value)*(standard error of b1)
Confidence Interval =  15.0352 ± (1.681071)*(0.4831)
  15.0352 + ((1.681071)*(0.4831))
  15.0352 - ((1.681071)*(0.4831))

Confidence Interval = (14.22307, 15.84733)


# b. Conduct a t test to determine whether or not there is a linear association between X and Y here; control the a risk at .10. State the alternatives, decision rule, and conclusion. What is the P-value of your test? In part (b), clearly layout the hypothesis, compute the test-statistic using relevant formula, compute the p-value, and state your conclusion in the context of the problem. Then, use the software output to confirm your results. (Similar requirement to part a)

```{r}
summary(cm.SLR) #in order to find estimated parameter & standard error of slope
tobs <- (15.0352 - 0)/(0.4831) # confirmed by summary above


```
H0: beta1 = 0
Ha: beta ≠ 0
T-test statistic = tobs = (b1 - 0)/(standard error of b1)
T-test statistic = tobs = (15.0352 - 0)/(0.4831) = 31.12233

Since the alternative hypothesis is two sided, the p-value is calculated as: 
p value = 2 * P(t >|tobs|), where t~t(n-2)

2*pt(31.12233,43) this is totally wrong and not sure what this written out formula wise should be


#c. Are your results in parts (a) and (b) consistent? Explain.

```{r}

```


# d. The manufacturer has suggested that the mean required time should not increase by more than 14 minutes for each additional copier that is serviced on a service call. Conduct a test to decide whether this standard is being satisfied by Tri-City. Control the risk of a Type I error at .05. State the alternatives, decision rule, and conclusion. What is the P-value of the test? 
# In part (d), test Ha: beta1 > 14. (Why?) Clearly state the hypothesis, use software to get necessary intermediate results, compute the test-statistic using relevant formula, compute the p-value, and state your conclusion in the context of the problem


```{r}


```
 
## Question 4 (Problem 2.14 a, b)
# In both part a and b, compute the confidence intervals using relevant formula(s). Follow these steps when “formula-based” computation is requested. First, use software to get the intermediate results, such as the  estimated valuesand their standard errors. Next, plug the intermediate results into the appropriate formula to construct the confidence interval or prediction interval. Then, use software directly to confirm your results. In SPSS, you can save the CI or PI for the predicted y-values to the data table. In R, use function predict() to compute CI or PI for predicted y-values.

# a. 
```{r}

```

# b.
```{r}

```

## Question 5 
Review the example in Lab 1. R users see item 8 in RLab1_SLR.pdf. SPSS users see
p.6 of Lab1_SLR.pdf. Practice drawing the “confidence interval band” and “prediction
interval band” on the scatter plot with the regression line. Turn in a graph using Copier
Maintenance data
```{r}

```

## Question 6 (Problem 2.24 (b, c, d)) 
Hint: Refer to the ANOVA table in the output.

# a. 
```{r}

```

#b.In part b, identify the relevant Sum of Squares from the output. Then verify the computation of the F-test, which is already in the output. Clearly layout the hypothesis, compute the test-statistic using relevant formula, compute the p-value, and statement your conclusion in the context of the problem.
```{r}

```

# Question 7
Refer to the Copier Maintenance data. Learn to read basic output of regression analysis
from another software. We will use SPSS output in the midterm exam.

# a. R users: Review handout Reading_SPSS_output.pdf. Compare the SPSS output to the R output in Homework 1, 2 and 3, and match the result. 

I have learned how to read basic output for regression analysis in R and SPSS.


