---
title: "HW2OliviaD'Souza"
author: "Olivia D'Souza"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 
Work on Lab 1 if you haven’t done so already. Be sure to read the instructions and the
sample submissions carefully before you start. You are expected to reproduce most of
the outputs in Lab 1 by the end of class on Tuesday, Sept 12.

## Question 2
(This will be in Quiz.) Please review the concept of computing p-values 
and think about how to find the p-value of a one-sided (aka. one-tailed) test from t
he p-value of a two-sided (aka. two-tailed) test. Here are some references:

The following questions use the Copier Maintenance data from Problem 1.20 (p.24)
## Question 3: Problem 2.5 (a, b, c, d) (2 pts)

```{r}
#read Copier Maintenance data in
cm <- read.table("ch2dataset.txt", header=F)
colnames(cm) <- c("time", "copiers")
head(cm, 3)
tail(cm, 2)
```


# a. Estimate the change in the mean service time when the number of copiers serviced increases by one. Use a 90 percent confidence interval. Interpret your confidence interval.Part (a) is asking about the slope. (Why?) 


```{r}

cm.SLR<-lm(time~copiers, data=cm). #estimated linear regression
cm.SLR
summary(cm.SLR) # In order to find the estimated parameter & standard error
critVal <-qt(.95,43) #finds critcal value based on 90% confidence int
critVal
confint(cm.SLR, level=0.90) # confirms my math using R software


```

Part A is asking about the slope because it asks for the "change in mean service time (y variable)
when the number of copiers serviced (x variable) increases by one. This is the definition
of the slope since you interpret the value as the change in mean y for one unit increase of x. 

Confidence Interval = est. of b1 ± (critical value)*(standard error of b1)
Confidence Interval =  15.0352 ± (1.681071)*(0.4831)
  15.0352 + ((1.681071)*(0.4831))
  15.0352 - ((1.681071)*(0.4831))

Confidence Interval = (14.22307, 15.84733)


# b. Conduct a t test to determine whether or not there is a linear association between X and Y here; control the a risk at .10. State the alternatives, decision rule, and conclusion. What is the P-value of your test?

```{r}

```



#c. Are your results in parts (a) and (b) consistent? Explain.

# d. The manufacturer has suggested that the mean required time should not increase by more than 14 minutes for each additional copier that is serviced on a service call. Conduct a test to decide whether this standard is being satisfied by Tri-City. Control the risk of a Type I error at .05. State the alternatives, decision rule, and conclusion. What is the P-value of the test?






```{r}


```



```{r pressure, echo=FALSE}
plot(pressure)
```


